### 2024-07-12 analysis/preprocessing.qmd
created classification preprocessing and analysis parts

```{r}
#| warning: false
# get metaMER df:
meta_df <- get_metaMER_df(path_2_studies = here::here('studies'))

@@

@@

print(paste("Where", length(unique(metaMER_results$citekey[metaMER_results$model_category=='classification'])), "are classification studies"))
# note that we have some classification studies that also do regression and vice versa?

@@

```{r}
R_studies <- dplyr::filter(metaMER_results,model_category=='regression' & str_detect(measure,'r|cc|r2|R2')) #  
dim(R_studies)

# eliminate measures that we don't need now
R_studies <- dplyr::filter(R_studies,!str_detect(statistic,'r2-95l|r2-95u|pvalue|upper|lower|sd|null|rmse'))
R_studies <- dplyr::filter(R_studies,!str_detect(measure,'nFeatures|rmse|jensen shannon divergence|accuracy|ccc|vector distance'))

table(R_studies$measure)

@@

```{r}
R_studies$dimension[str_detect(R_studies$dimension,'activation|energy arousal|tension arousal')]<-'arousal'
R_studies$dimension[str_detect(R_studies$dimension,'pleasantness')]<-'valence'
R_studies <- dplyr::filter(R_studies,!str_detect(dimension,'av')) # relates to distances, can be omitted

@@

# Deal with four studies involving multiple datasets: 
R_studies$stimulus_n[R_studies$stimulus_n==" emoMusic: 1000, soundtracks: 360, chinese: 500 "] <- 938 #resolved from the paper
R_studies$stimulus_n[R_studies$stimulus_n==" 2372 (subset of PSIC3839, total n: 3839)    "] <- 2372 # resolved
R_studies$stimulus_n[R_studies$stimulus_n==" study 1: 20; study 2: 40) % three outliers  "] <- 40 # decided to take this from validation
# REDO with a clearer function
eliminate <- str_detect(R_studies$unique_id,"hu2017cr") & !str_detect(R_studies$unique_id,"all")
R_studies <- R_studies[!eliminate,]
R_studies$stimulus_n[R_studies$stimulus_n==" MER60: 60, CH818: 818, AMG1608: 1608  "] <- 60+818+1608 #
R_studies$stimulus_n <- as.numeric(R_studies$stimulus_n)

@@

@@

R_summary <- summarise(group_by(R_studies,dimension,citekey),valuesMean=mean(values,na.rm=T),valuesMedian=median(values,na.rm=T),valuesMax=max(values,na.rm=T),stimulus_n=first(stimulus_n),studyREF=first(studyREF),model_class_id=first(model_class_id))

@@

```{r}
C_studies <- dplyr::filter(metaMER_results,model_category=='classification' & str_detect(measure,'accuracy|r2|mcc')) #  
dim(C_studies)

# eliminate measures that we don't need now
C_studies <- dplyr::filter(C_studies,values>0)
C_studies$statistic[is.na(C_studies$statistic)]<-'Not defined'
C_studies <- dplyr::filter(C_studies,!str_detect(statistic,'pvalue|upper|lower'))
C_studies <- dplyr::filter(C_studies,str_detect(measure,'accuracy|mcc|r2'))
dim(C_studies)

table(C_studies$measure,C_studies$statistic)

dim(C_studies)
# Conversion into mcc (==r), do we need

```
## Homogenise the outcome variable names to valence and arousal

@@

# Deal with four studies involving multiple datasets: 
C_studies$stimulus_n[C_studies$stimulus_n==" 429; 350 popular songs + 79 songs from the Beatles (Mirex 2009 collection)  "] <- 429 # resolved from the paper
C_studies$stimulus_n[C_studies$stimulus_n==" 5192; 12 per user in user validation (not included here due to little information),   AcousticBrainz validation: 60000  "] <- 5192 # resolved
C_studies$stimulus_n[C_studies$stimulus_n==" ISMIR2012: 2886, NJU_V1: 777, Hindi: 1037  "] <- 2886+777+1037 # decided to take this from validation
C_studies<-C_studies[!str_detect(C_studies$stimulus_n,'not specified'),]
C_studies<-C_studies[!str_detect(C_studies$stimulus_n,'Specified elsewhere'),]
C_studies$stimulus_n <- as.numeric(C_studies$stimulus_n)

@@

g1<-ggplot(C_studies,aes(x=values,fill=citekey,color=dimension))+
  geom_histogram()+
  facet_wrap(.~model_class_id)+
  scale_color_manual(values = c('black','white'))+
  theme_dark()+
  scale_x_continuous(breaks = seq(0,1,by=.1))

@@

R_summary <- summarise(group_by(R_studies,dimension,citekey),valuesMean=mean(values,na.rm=T),valuesMedian=median(values,na.rm=T),valuesMax=max(values,na.rm=T),stimulus_n=first(stimulus_n),studyREF=first(studyREF),model_class_id=first(model_class_id))

@@

@@

@@

g3 <- ggplot(R_studies,aes(x=year,y=values,colour=model_class_id))+
  geom_point(show.legend = T)+
  facet_wrap(.~dimension)+
  theme_bw()
g3
```
## Simple model complexity metric

@@

@@

@@

@@

```{r}
#write.csv(x = R_studies,file = 'R_studies.csv')
#write.csv(x = R_summary,file = 'R_summary.csv')
```

>>>>

```{r}
#| warning: false
library(stringr) 
# get metaMER df:
meta_df <- get_metaMER_df(path_2_studies = here::here('studies'))

@@

### Classify stimulus genres

```{r}
#| eval: true
metaMER_results$stimulus_genre_mixed <- 'SingleGenre'
metaMER_results$stimulus_genre_mixed[str_detect(metaMER_results$stimulus_genre,',|multi')] <- 'MultiGenre' # Class name from Elements of Stat.."
#table(metaMER_results$stimulus_genre_mixed)
```

### Classify journals

```{r}
metaMER_results$journal_type <- "Engineering"
metaMER_results$journal_type[str_detect(metaMER_results$journal,'Quarterly Journal of Experimental Psychology|PSYCHOLOGY OF MUSIC|PLOS ONE|JOURNAL OF NEW MUSIC RESEARCH|IEEE-ACM TRANSACTIONS ON AUDIO SPEECH AND LANGUAGE PROCESSING|FRONTIERS IN PSYCHOLOGY|Frontiers in Psychology')] <- 'Psychology' # Class name from Elements of Stat.."

#table(metaMER_results$journal_type)
```

@@

print(paste("Where", length(unique(metaMER_results$citekey[metaMER_results$model_category=='classification'])), "are classification studies"))
# note that we have some classification studies that also do regression and vice versa?
# THIS IS CORRECT:
# [1] "We have 1192 observations"
# [1] "We have 37 studies"
# [1] "Where 25 are regression studies"
# [1] "Where 12 are classification studies"

@@

```{r}

unique(metaMER_results$citekey[metaMER_results$model_category=="regression"])

R_studies <- dplyr::filter(metaMER_results, measure %in% c('r','ccc','pcc','cc','r2','R2')) #  
R_studies$statistic
dim(R_studies)

# explicit NA coding
R_studies$statistic[is.na(R_studies$statistic)]<-'r2'

# include measures we need 
R_studies <- dplyr::filter(R_studies,str_detect(statistic,'adjusted|ccc|pcc|cc|globalOptimal|localOptimal|mean|r2|r'))
dim(R_studies)
table(R_studies$measure)

@@

```{r}
R_studies$dimension[str_detect(R_studies$dimension,'activation|energy arousal|tension arousal')]<-'arousal'
R_studies$dimension[str_detect(R_studies$dimension,'pleasantness')]<-'valence'
R_studies <- dplyr::filter(R_studies,!str_detect(dimension,'av|resonance')) # relates to distances, can be omitted

@@

# Deal with four studies involving multiple datasets: 
R_studies$stimulus_n[R_studies$stimulus_n==" emoMusic: 744, soundtracks: 360, chinese: 500 "] <- 938 # resolved from the paper
R_studies$stimulus_n[R_studies$stimulus_n==" 2372 (subset of PSIC3839, total n: 3839)    "] <- 2372 # resolved
R_studies$stimulus_n[R_studies$stimulus_n==" study 1: 20; study 2: 40) % three outliers  "] <- 40 # decided to take this from validation
# REDO with a clearer function
eliminate <- str_detect(R_studies$unique_id,"hu2017cr") & !str_detect(R_studies$unique_id,"all")
R_studies <- R_studies[!eliminate,]
R_studies$stimulus_n[R_studies$stimulus_n==" MER60: 60, CH818: 818, AMG1608: 1608  "] <- 60+818+1608 #
R_studies$stimulus_n <- as.numeric(R_studies$stimulus_n)

@@

g1
```
## Check studies included

```{r}
#| eval: false
included_in_summary <- R_summary$citekey[R_summary$dimension=='arousal']

included_originally <- unique(metaMER_results$citekey[metaMER_results$model_category=='regression'])

included_originally[!included_originally %in% included_in_summary]

@@

R_summary <- summarise(group_by(R_studies,dimension,citekey),valuesMean=mean(values,na.rm=TRUE),valuesMedian=median(values,na.rm=TRUE),valuesMax=max(values,na.rm=TRUE),stimulus_n=first(stimulus_n),studyREF=first(studyREF),model_class_id=first(model_class_id),journal_type=first(journal_type),stimulus_genre_mixed=first(stimulus_genre_mixed))

@@

```{r}
C_studies<-metaMER_results
sum(is.na(C_studies$measure)) # 0
sum(is.na(C_studies$statistic)) # 426

C_studies$statistic[is.na(C_studies$statistic)]<-'acc'

C_studies <- dplyr::filter(C_studies, str_detect(measure,'mcc|accuracy|auc|classification|f1|fscore|fvalue|kappa|precision|recall')) #
table(C_studies$measure)
dim(C_studies)
table(C_studies$statistic)

C_studies <- dplyr::filter(C_studies, str_detect(statistic,'mean|median|acc')) #  
dim(C_studies)
length(unique(C_studies$citekey))

table(C_studies$measure,C_studies$statistic)
table(C_studies$measure,C_studies$citekey)
table(C_studies$statistic,C_studies$citekey)

# ToDo
# how to convert accuracy (precision, recall, f) into kappa or mcc
# There IS NO solution unfortunately, accuracy and mcc can vary independently, but
# see Chicco and Jurman, 2020, https://doi.org/10.1186/s12864-019-6413-7
# which shows that MCC is the best. The paer compares accuracy and F1 to MCC 
# and shows high correlations between accuracy and MCC (r=0.90-0.92). 
# It is only the exceptions (unbalanced datasets with binary classification) that deliver major discrepancies. Our dataset does not seem to be having these, so
# perhaps we can equate accuracy with MCC in a minority of cases.

# Conversion
# convert one 'classification' into accuracy (bai2017mu)
# convert one 'f1' into accuracy (hu2022de)
C_studies$measure[C_studies$citekey=='bai2017mu'] <- 'accuracy'
C_studies$measure[C_studies$citekey=='hu2022de' & C_studies$measure=='f1']<-'accuracy'
dim(C_studies)
C_studies <- dplyr::filter(C_studies, str_detect(measure,'mcc|accuracy')) #  
dim(C_studies)
table(C_studies$measure,C_studies$citekey)

# take accuracy as MCC in those studies that have not specified mcc
C_studies$accuracy_converted_to_mcc <- 0
C_studies$accuracy_converted_to_mcc[C_studies$citekey=='bai2016di' | 
                  C_studies$citekey=='bai2017mu' | 
                  C_studies$citekey=='hizlisoy2021mu' |
                  C_studies$citekey=='hu2022de' |    
                  C_studies$citekey=='yeh2014po' |
                  C_studies$citekey=='zhang2016br' |
                  C_studies$citekey=='zhang2017fe' 
                  & C_studies$measure=='accuracy']<-1

C_studies$measure[C_studies$citekey=='bai2016di' | 
                  C_studies$citekey=='bai2017mu' | 
                  C_studies$citekey=='hizlisoy2021mu' |
                  C_studies$citekey=='hu2022de' |    
                  C_studies$citekey=='yeh2014po' |
                  C_studies$citekey=='zhang2016br' |
                  C_studies$citekey=='zhang2017fe' 
                  & C_studies$measure=='accuracy']<-'mcc'
table(C_studies$measure,C_studies$citekey)
C_studies <- dplyr::filter(C_studies, str_detect(measure,'mcc')) #  
dim(C_studies)
table(C_studies$measure,C_studies$accuracy_converted_to_mcc)
t.test(C_studies$values,y = C_studies$accuracy_converted_to_mcc) # There is a difference

```
## Homogenise the stimulus N

@@

# Deal with four studies involving multiple datasets: 
C_studies$stimulus_n[C_studies$stimulus_n==" 429; 350 popular songs + 79 songs from the Beatles (Mirex 2009 collection)  "] <- 429 # resolved from the paper
C_studies$stimulus_n[C_studies$stimulus_n==" 5192; 12 per user in user validation (not included here due to little information),   AcousticBrainz validation: 60000  "] <- 5192 # resolved
C_studies$stimulus_n[C_studies$stimulus_n==" ISMIR2012: 2886, NJU_V1: 777, Hindi: 1037  "] <- 2886+777+1037 # decided to take this from validation
C_studies$stimulus_n[C_studies$stimulus_n==" total: 564; unambiguous: 416, circular validation: 39 "] <- 564 # 
C_studies$stimulus_n <- as.numeric(C_studies$stimulus_n)

@@

```{r}
#| eval: false
library(ggplot2)

g1<-ggplot(C_studies,aes(x=values,fill=citekey,color=dimension))+
  geom_histogram()+
  facet_wrap(.~model_class_id)+
  theme_dark()+
  scale_x_continuous(breaks = seq(0,1,by=.1))

@@

- [ ] Ask Cameron whether it is OK to treat dimension contents as they are all classification studies (the exceptions are `av` and `valence` and `arousal`).

```{r}
library(tidyverse)
C_studies$citekey <- factor(C_studies$citekey)

C_summary <- summarise(group_by(C_studies,citekey),valuesMean=mean(values,na.rm=TRUE),valuesMedian=median(values,na.rm=TRUE),valuesMax=max(values,na.rm=TRUE),stimulus_n=first(stimulus_n),studyREF=first(studyREF),model_class_id=first(model_class_id), stimulus_genre_mixed=first(stimulus_genre_mixed),journal_type = first(journal_type))

@@

## Visualise Summary

```{r}

g2 <- ggplot(C_summary,aes(x=stimulus_n,y=valuesMax,label=citekey,color=stimulus_genre_mixed))+
  geom_point()+
  geom_label_repel(size=1.5)+
#  coord_flip()+
  theme_bw()
g2

g3 <- ggplot(C_summary,aes(x=stimulus_n,y=valuesMax,label=citekey,color=model_class_id))+
  geom_point()+
  geom_label_repel(size=1.2)+
#  coord_flip()+
  theme_bw()
g3

## could be more informative when done with the full data
C_studies$citekey<-factor(C_studies$citekey)

@@

g4 <- ggplot(C_studies,aes(x=stimulus_n,y=values,label=citekey,color=model_class_id,fill=model_class_id))+
  geom_point(size=4)+
  geom_label_repel(size=1.2, max.overlaps=50,show.legend = T,color='white')+
#  scale_x_continuous(breaks = seq(0,1,by=.25),limits = c(0,1))+
#  scale_y_continuous(breaks = seq(0,1,by=.25),limits = c(0,1))+
  theme_bw()
g4

@@

## Simple model complexity metric based on feature_n

```{r}
table(C_studies$feature_n)
C_studies$feature_n[str_detect(C_studies$feature_n,'eight different ')]<-600 # arbitrary!
C_studies$feature_n[str_detect(C_studies$feature_n,'126, retained 97')]<-126 
C_studies$feature_n[str_detect(C_studies$feature_n,'1702; best model uses 100 after reduction')]<-1702 #  arbitrary!
C_studies$feature_n[str_detect(C_studies$feature_n,'548; after reduction, ')]<-548 
C_studies$feature_n[str_detect(C_studies$feature_n,'548. Post-reduction')]<-548 
C_studies$feature_n[str_detect(C_studies$feature_n,'8; 3 after')]<-8 
C_studies$feature_n[str_detect(C_studies$feature_n,'between 9 and 10')]<-10 
C_studies$feature_n[str_detect(C_studies$feature_n,'summarize feature')]<-600 # arbitrary
table(C_studies$feature_n)

C_studies$feature_n<-as.numeric(C_studies$feature_n)

print(quantile(C_studies$feature_n,c(0.333,0.666),na.rm = TRUE))
#quantile(R_studies$feature_n,c(0.25,0.50,0.75),na.rm = T)

# Assign
C_studies$feature_n_complexity <- cut(C_studies$feature_n,
                                      breaks = c(0,
                                                 as.numeric(quantile(C_studies$feature_n,c(0.333),na.rm = TRUE)),
                                                 as.numeric(quantile(C_studies$feature_n,c(0.666),na.rm = TRUE)),
                                                 1000),
                                      labels = c("Feature n < 119","Feature n > 119 & < 548","Feature n > 548"))
table(C_studies$feature_n_complexity)
```

@@

Needs to be done from the unsummarised data (`C_studies`).

```{r}
#| eval: true
tmp <- drop_na(C_studies)
library(ggdist)

@@

g <- ggplot(tmp,aes(x=model_class_id,y=values,color=citekey,label=citekey))+
  stat_halfeye(aes(fill=citekey),point_interval="mean_qi", trim=FALSE, expand=FALSE, show.legend = FALSE,adjust = 1.25, density="bounded", point_size=3,scale = 1,alpha=0.5) + 
  geom_point(alpha=0.5,show.legend = F,position = position_jitter(width = .3))+
  #geom_label_repel(size=2,max.overlaps = 50)+
  facet_wrap(.~feature_n_complexity)+
  ylab("Correlation Coefficient")+
  xlab("Model Technique")+
  scale_y_continuous(limits = c(0,1),expand = c(0.01,0.01))+
  geom_text_repel(aes(x = model_class_id, y = values, label = studyREF),
             stat = "summary", fun = mean,show.legend = F)+
  theme_bw()
print(g)
#ggsave(filename = 'FeatureN_regression.pdf',g,height = 7,width = 11)
```

@@

```{r}
TC <- NULL
TC$study_n <- length(unique(C_studies$citekey))
TC$model_n <- nrow(C_studies)
t<-table(C_studies$model_class_id)
t2 <- paste0(names(t),': ', as.numeric(t))
TC$model_types_n <- str_c(t2,collapse = "\n")
TC$feature_Desc <- paste0('Min=',min(C_studies$feature_n,na.rm = TRUE),', Md=',median(C_studies$feature_n,na.rm = TRUE),', Max=', max(C_studies$feature_n,na.rm = TRUE))
TC$stimulus_Desc <- paste0('Min=',min(C_studies$stimulus_n,na.rm = TRUE),', Md=',median(C_studies$stimulus_n,na.rm = TRUE),', Max=', max(C_studies$stimulus_n,na.rm = TRUE))
print(TC)

@@

```{r}
write.csv(x = C_studies,file = 'C_studies.csv')
write.csv(x = C_summary,file = 'C_summary.csv')
```

### 2024-07-12 manuscript/manuscript.qmd
created classification preprocessing and analysis parts

>>>>

| Model       | Models, obs|$r$ [95%-CI]             | $t$ | $p$  |$\tau^2$|$I^2$|
|:------------|:-----------|:------------------------|:----|:-----|:-----|:------|
| All         | 89,87347   | 0.8074 [0.7681; 0.8407] |21.4 |0.0001|0.2415|99.7%  |
| All Trim    | 29,6499    | 0.8185 [0.8046; 0.8314] | 58.3|0.0001|0.0066| 60.4%  |
| Max         | 14,17184   | 0.8564 [0.7386; 0.9234] | 10.3|0.0001|0.329 | 99.8% |
| Max Trim    | 6,3653     | 0.8689 [0.7760; 0.9249] | 11.6|0.0001|0.0749| 97.5% |
